---
title: "Accuracy, sensitivity, specificity, and precision"
author: "Rob Walker"
output: html_document
---

# Unbalanced targets
- smoke alarms
- airport security
- fraud detection
- uncontacted indigenous villages
- cancer diagnosis (e.g., endoscopy to find bowel cancer)

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(reshape2)

x <- matrix(c(20,180,10,1820), byrow = TRUE, 2, 2, dimnames=list(c("test positive","test negative"),c("cancer 1.5%","no cancer 98.5%")))
x

round(prop.table(x),3)*100 #percents

TP = 20
FN = 10
FP = 180
TN = 1820

(accuracy = (TP + TN) / (TP + TN + FP + FN))
(sensitivity = TP/(TP + FN)) # also called Recall
(specificity = TN/(TN + FP)) 
(precision = TP/(TP + FP)) # posterior probability of having cancer given a positive test

```

# Find optimal threshold
- find optimal k threshold in FICO credit scores at which to identify default loans
- real FICO credit scores and actual good/default data

```{r}
FICO = c(608,745,443,443,443,546,546,650,617,694,694,650,685,587,590,665,644,593,635,545,466,466,466, 720,646,569,569,619,573,582,582,643,664,664,560,560,600,600,532,614,532,434,468,616,489,680,462,462)

GoodOrDefault = c("Bad","Good","Good","Good","Bad","Good","Good","Bad","Good","Good","Good","Bad","Good","Good","Good","Good","Good","Good","Good","Good","Bad","Good","Bad","Bad","Bad","Good","Bad","Bad","Bad","Good","Good","Bad","Bad","Good","Good","Good","Good","Good","Bad","Good","Bad","Bad","Bad","Bad","Bad","Good","Good","Good")

df = data.frame(GoodOrDefault,FICO)
ggplot(df, aes(x=FICO, y=as.numeric(GoodOrDefault)-1)) + geom_jitter(height=0.05) +
  geom_smooth()

table(df$GoodOrDefault)

possible_k = seq(min(df$FICO),max(df$FICO), length.out = 1000)
accuracy <- sapply(possible_k, function(k) {
  predicted_class <- as.numeric(df$FICO < k)
  (sum(predicted_class == 1 & df$GoodOrDefault == "Bad") + sum(predicted_class == 0 & df$GoodOrDefault == "Good")) / dim(df)[1]
})
sensitivity <- sapply(possible_k, function(k) {
  predicted_class <- as.numeric(df$FICO < k) # predic default if FICO lower than threshold
  sum(predicted_class == 1 & df$GoodOrDefault == "Bad") / table(df$GoodOrDefault)[1]
})
specificity <- sapply(possible_k, function(k) {
  predicted_class <- as.numeric(df$FICO > k)
  sum(predicted_class == 1 & df$GoodOrDefault == "Good") / table(df$GoodOrDefault)[2]
})
precision <- sapply(possible_k, function(k) {
  predicted_class <- as.numeric(df$FICO < k)
  sum(predicted_class == 1 & df$GoodOrDefault == "Bad")/ sum(predicted_class)
})
out = data.frame(FICO=possible_k,specificity,sensitivity,precision,accuracy)
out <- melt(out, id.vars = "FICO", variable.name = "Metric", value.name = "Data")

ggplot(out, aes(x = FICO, y = Data, color = Metric)) + geom_line(size=2) +
  ylab("") + xlab("FICO") + scale_color_manual(name="", labels=c("Specificity","Sensitivity","Precision","Accuracy"), values=c("red","green","blue","orange"))


```

# Summary
- Inherent tradeoff between sensitivity and specificity/precision
- Optimal threshold is a business decision, or a moral one
